---
title: 
output: 
  html_document:
    self_contained: false 
    highlight: tango
---

### 這份文件參照EP17-A2, Evaluation of Detection Capability for Clinical Laboratory Measurement Procedures; Approved Guideline- Second Editon實現Probit analysis，其中Data的部分在此文件的第51頁。

### Probit Analysis是用來找出檢測方法的LOD（Limit of Detection），通常都會設定為可偵測到95%陽性樣本量的最小濃度。若在用藥上，則是找出半致死劑量LD50（Lethal Dose, 50%）。此處處要以LOD為主。

### Code from: [http://dx.plos.org/10.1371/journal.pone.0054092.s004](http://dx.plos.org/10.1371/journal.pone.0054092.s004)。

## Sample Code

### 一開始會需要知道每個濃度對應到的樣本數，以及偵測到陽性的樣本數。

```{r}
posi=c(11,15,23,29,32,32,32)
total=c(30,30,32,32,32,32,32)
conce=c(0.006,0.014,0.025,0.05,0.15,0.3,0.5)

```



### 1. 利用General Linear Model，在family這個參數選擇`binomial(link="probit")`
  
###2. 在給變數的時候也要注意，實際上必須同時給偵測為陽性和陰性的數量，同時濃度必須取log（還蠻好理解的，probit有點像logistic）
###  $$ y=\beta _{1}x-\beta _{0} $$
  
###3. 得到 $\beta _{1}$ 和 $\beta _{0}$ 後，就可以將normal distribution上95%的quantile function帶入，再帶入10的指數函數，就可求得LoD。
###4. 這部分的信賴區間作法比較不熟，不多作解釋。


```{r}
LD <- function(r, n, d, conf.level) {
  #Set up a number series 
  p <- seq(1, 99, 1)
  
  #r=number responding, n=number treated, d=dose (untransformed), confidence interval level
  mod <- glm(cbind(r, (n-r)) ~ log10(d), family = binomial(link=probit))
  #Calculate heterogeneity correction to confidence intervals according to Finney, 1971
  #(p.72, eq. 4.27; also called "h")
  het = deviance(mod)/df.residual(mod)
  if(het < 1){het = 1} 
  
  #Heterogeneity cannot be less than 1
  
  #Extract slope and intercept
  summary <- summary(mod, dispersion=het, cor = F)
  intercept <- summary$coefficients[1]
  interceptSE <- summary$coefficients[3]
  slope <- summary$coefficients[2]
  slopeSE <- summary$coefficients[4]
  z.value <- summary$coefficients[6]
  N <- sum(n)
  
  #Intercept (alpha)
  b0<-intercept
  #Slope (beta)
  b1<-slope
  #Slope variance 
  vcov = summary(mod)$cov.unscaled
  var.b0<-vcov[1,1]
  #Intercept variance
  var.b1<-vcov[2,2]
  #Slope intercept covariance
  cov.b0.b1<-vcov[1,2]
  
  #Adjust alpha depending on heterogeneity (Finney, 1971, p. 76)
  alpha<-1-conf.level
  if(het > 1) {talpha <- -qt(alpha/2, df=df.residual(mod))} else {talpha <- -qnorm(alpha/2)}
  
  #Calculate g (Finney, 1971, p 78, eq. 4.36)  
  #"With almost all good sets of data, g will be substantially smaller than 1.0 and seldom greater than 0.4."
  g <- het * ((talpha^2 * var.b1)/b1^2)
  
  #Calculate theta.hat for all LD levels based on probits in eta 
  #(Robertson et al., 2007, pg. 27; or "m" in Finney, 1971, p. 78)
  eta = family(mod)$linkfun(p/100)  #probit distribution curve
  theta.hat <- (eta - b0)/b1
  
  #Calculate correction of fiducial limits according to Fieller method 
  #(Finney, 1971, p. 78-79. eq. 4.35) 
  const1 <- (g/(1-g))*(theta.hat + cov.b0.b1/var.b1) 
  const2a <- var.b0 + 2*cov.b0.b1*theta.hat + var.b1*theta.hat^2 - g*(var.b0 - (cov.b0.b1^2/var.b1))
  const2 <- talpha/((1-g)*b1) * sqrt(het * (const2a))
  
  #Calculate the confidence intervals LCL=lower, UCL=upper 
  #(Finney, 1971, p. 78-79. eq. 4.35) 
  LCL <- (theta.hat + const1 - const2)
  UCL <- (theta.hat + const1 + const2)
  
  #Calculate variance for theta.hat 
  #(Robertson et al., 2007, pg. 27)
  var.theta.hat <- (1/(theta.hat^2)) * ( var.b0 + 2*cov.b0.b1*theta.hat + var.b1*theta.hat^2 )
  
  #Make a data frame from the data at all the different values
  ECtable <- data.frame(
    "probability"=p,
    "N"=N,
    "EC"=10^theta.hat,
    "LCL"=10^LCL,
    "UCL"=10^UCL, 
    "slope"=slope, 
    "slopeSE"=slopeSE, 
    "intercept"=intercept, 
    "interceptSE"=interceptSE, 
    "z.value"=z.value, 
    "chisquare"=deviance(mod), 
    "df"=df.residual(mod), 
    "h"=het, 
    "g"=g,
    "theta.hat"=theta.hat,
    "var.theta.hat"=var.theta.hat)
  
  #Select output level
  return(ECtable)
}
conf.level=0.95
LoD.level=95
test <- LD(posi,total,conce,conf.level)
print(paste("LoD", "= ", round(test[LoD.level,"EC"], 3) , ',',conf.level*100, "% CI= ",round(test[LoD.level,"LCL"], 3),"-", round(test[LoD.level,"UCL"], 3)))
```

### 畫圖表示不同Probability與濃度之間的關係，並加上虛線表示95%信賴區間
```{r}
library(ggplot2)
ggplot(test, aes(y=probability))+
  geom_line(aes(x=EC))+
  geom_line(aes(x=LCL), linetype=2)+
  geom_line(aes(x=UCL), linetype=2)+
  geom_hline(yintercept = 95, color='red')+
  scale_x_log10()
```

